Sensitivity Analysis

### 2024-06-07 Article_Sensitivity_Analsysis_short-Chamomile.tex
07.06

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
The article presents a comprehensive study on the supercritical fluid extraction of essential oil from chamomile flowers, focusing on model dynamics and applying a distributed-parameter model to describe the fluid-solid extraction process. By employing the concept of quasi-one-dimensional flow, the study simplifies the spatial dimensions of the extraction process, ensuring uniform flow across any cross-section while allowing for variations in the area available for the fluid phase. The physical properties of the solvent are estimated using the Peng-Robinson equation of state. The model was validated under multiple constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. The model dynamic is investigated by analysing how sensitive the model output is to model parameters. In this work, the local sensitivity analysis is introduced and used to examine the influence of infinitely small changes in the pressure on the model's state space and the extraction yield.

@@

\subsection{Sensitivity Analysis} \label{CH: Sensitivity_Analysis}
\subfile{Sections/Sensitivity_Analysis}

@@

Sensitivity analysis is a tool to understand how parameters affect a model's output. In the case of dynamical systems, local sensitivity analysis provides a time series describing how that dependency evolves. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. The local sensitivity analysis techniques consider only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear model. The sensitivity equations can be obtained in various ways. This work implemented the automatic differentiation technique to get the sensitivity equations, which evaluate the effect of pressure increase on the supercritical extraction model. At given operating conditions ($35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s), the step change of pressure enhance the mass transfer, which leads to faster loss of solute from solid particles and consequently to negative sensitivities in the solid phase. An analogous response is visible in the fluid phase, where the concentration of solute increases and is visible as positive sensitivities. As more solute is transported to the fluid phase, the extraction yield is improved and characterised by positive sensitivities. Local sensitivity analysis can provide valuable information about process modelling, experiment design, or model reduction by identifying which parameters are influential and how these influence changes over time.

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study aimed to investigate the supercritical extraction process of caraway oil from chamomile flowers. The distributed-parameter model describes the fluid-solid extraction process. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. A set of laboratory experiments was performed under multiple constant operating conditions: $30 - 40^\circ C$, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. Sensitivity analyses play a crucial role in assessing the robustness of the findings or conclusions based on mathematical model. The local sensitivity analysis investigates the influence of infinitely small changes in the inlet temperature, pressure, and flow rate on the extraction yield.

@@

\subsection{Local sensitivity analysis} \label{CH: Sensitivity_Analysis}
\subfile{Sections/Sensitivity_Analysis}

@@

Sensitivity analysis is a tool to understand how parameters affect a model's output. In the case of dynamical systems, local sensitivity analysis provides a time series describing how that dependency evolves. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. The local sensitivity analysis techniques consider only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear model. The sensitivity equations can be obtained in various ways. This work implemented the automatic differentiation technique to derive the sensitivity equations. The effect of pressure increase on the outcome of the supercritical extraction model is analysed in this work. At given operating conditions ($35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s), the step change of pressure enhance the mass transfer, which leads to faster loss of solute from solid particles and consequently to negative sensitivities in the solid phase.  Analogously, the sensitivities in the fluid phase are characterized by positive deviations, which indicates that more solute is transported to the fluid phase. As a result, the extraction yield is improved and characterised by positive sensitivities as well. Local sensitivity analysis can provide valuable information about process modelling, experiment design, or model reduction by identifying which parameters are influential and how these influence changes over time.

### 2024-06-07 Sections/Literature_Review.tex
07.06

	\citet{Santos2000}, in their work, considered a model of supercritical extraction process for semi-continuous isothermal and isobaric extraction process using carbon dioxide as a solvent. The parametric sensitivity analysis was carried out by applying disturbances of 10\% in the values of the normal operation conditions.

	\citet{Hatami2024} used a one-factor-at-a-time sensitivity analysis to assess the response of NPV concerning variations in both technical and economic variables. Their findings show that the most influential factors on NPV include the price of the extract, the interest rate, the dynamic time of SFE, and the project lifetime.

	\citet{Poletto1996} provided a general dimensionless model for the supercritical extraction process of vegetable and essential oils and applied a sensitivity analysis. They found that a dimensionless partition coefficient and a dimensionless characteristic time appeared as the most important parameters of the extraction process. The sensitivity calculations were performed by varying the parameters and analysing the model response.

>>>>

	\citet{Santos2000}, in their work, considered the model of \citet{Sovova1994} for the semi-continuous isothermal and isobaric extraction processes using carbon dioxide as a solvent. The parametric sensitivity analysis using a factorial design in two levels was carried out. The model parameters were disturbed by 10\%, and their main effects were analysed so that it is possible to propose strategies for high-performance operation. The authors performed the sensitivity analysis with respect to the superficial velocity, particle diameter, initial concentration of solute in the solid phase and the concentration of solute in the fluid phase at the inlet to the extractor.

	\citet{Hatami2024} performed a one-factor-at-a-time sensitivity analysis to	assess the response of net present value concerning variations in both technical and economic variables. The analysis is divided into two parts. In the first part, the sensitivity analysis is carried out for a SFE plant with a constant	extractor volume of 300 L. Here, the focus is on examining how net present value is influenced by changes in individual technical and economic parameters while keeping the extractor volume fixed. In the second part, the investigation revolves around the effects of altering the extractor volume (between 1 to 600 L) on the overall profitability of the project. Their findings show that the most influential factors on NPV include the price of the extract, the interest rate, the dynamic time of SFE, and the project lifetime.

	\citet{Poletto1996} provided a general dimensionless model for the supercritical extraction process of vegetable and essential oils and applied a sensitivity analysis. The sensitivity calculations were performed by varying the parameters and analysing the model response. The authors found that a dimensionless partition coefficient and a dimensionless characteristic time appeared as the most important parameters of the extraction process.

### 2024-06-07 Sections/Model.tex
07.06

	The fluid density $\rho_f$ can be obtained from an equation of state (Appendix \ref{subsubsec: Equation of state}) if the temperature and thermodynamic pressure are known along $z$. Variation in fluid density may occur due to pressure or inlet temperature changes. In a non-isothermal case, in Equations \ref{EQ:Velocity_1} and \ref{EQ:Velocity_2} $\rho_f$ is considered the average fluid density along the extraction column.

@@

			The heat balance equation describe the evolution of the internal energy in the system and it is given by Equation \ref{EQ:Enthalpy_equation}

@@

			The main advantage of this formulation is the presence of term $\partial P / \partial t $, which directly affects the system through the change of thermodynamic pressure (which is a control variable). %The enthalpy is related to the pressure and temperature through the following equation:

			If the value of enthalpy $h$ is known from the time evolution of the energy equation, and pressure $P$ is known from measurement, then the temperature $T$ can be reconstructed based on the departure function. The departure function is a mathematical function that characterizes the deviation of a thermodynamic property (enthalpy, entropy, and internal energy) of a real substance from that of an ideal gas at the same temperature and pressure. The departure function is defined as the difference between the value of a thermodynamic property for a real fluid and the corresponding value for an ideal gas at the same temperature and pressure. They are computed by integrating a function that depends on the equation of state and its derivatives. As presented by \citet{Gmehling2019}, for the Peng-Robinson equation of state, the enthalpy can be defined by Equation \ref{EQ:Enthalpy_PR}.

@@

			The pressure term in the energy equation, given by Equation \ref{EQ:Enthalpy_equation}, describes the change of the thermodynamic pressure with respect to time. As explained in Chapters \ref{CH:Governing_equations_chapter}, at Low-Mach number conditions, the thermodynamic pressure is nearly constant in space due to the small pressure wave propagation that occurs at the speed of sound. Under such conditions, the term $\partial P/\partial t$ can be approximated by a difference equation, which describes the pressure change in the whole system. The pressure ($P$) in the system is considered a state variable, while the pressure in the new time-step ($P_{in}$) is considered a control variable.

			{\footnotesize
				\begin{equation}
					\frac{\partial {\color{black}P}(t)}{\partial t} \approx \frac{{\color{black}P_{in}}(t) - {\color{black}P}(t) }{\Delta t}
			\end{equation}}

			Such a simplified equation allows for instantaneous pressure change in the system but does not consider a gradual pressure build-up and the effects of pressure losses. In a real system, the dynamics of pressure change would depend on a pump used in an extraction system and a back-pressure regulator used to control an outlet valve.

@@

			{\footnotesize
				\begin{align*}
					&c_f(t = 0, z) = 0  && c_s(t = 0, z) = c_{s0} && h(t = 0, z) = h_0 \\
					&c_f(t,   z=0) = 0  && h(t, z=0) = h_{in}(t)  && \frac{\partial c_f(t,z=L)}{\partial x} = 0 \\
					&\frac{\partial h(t,z=L)}{\partial x} = 0   && c_s(t, z=\{0,L\}) = 0 && y(0) = 0 \quad P(0) = P_0 \\
			\end{align*} }

@@

				{\footnotesize
					\begin{align*} \label{discretization}
						\dot{x}(t) &= \cfrac{d x}{d t} = 
						\begin{bmatrix}
							\cfrac{d c_{f,1}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{f,N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d c_{s,1}(t)}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{s,N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d h_1(t)}{d t} 	  \\
							\vdots 					  \\
							\cfrac{d h_{N_z}(t)}{d t} \\
							\\ \hline \\
							\cfrac{d P(t)}{d t} \\
							\\ \hline \\
							\cfrac{d y(t)}{d t}
						\end{bmatrix}
						=
						\underbrace{\begin{bmatrix}
								G_1 \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\ 
								G_{N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \hline \\ \\
								G_{N_z+1} \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\
								G_{2N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \\ \hline \\ 
								G_{2N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\vdots\\
								G_{3N_z} \left( c_f,c_s,h; \Theta \right) \\ 
								\\ \\ \hline \\
								G_{3N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\\ \\ \hline \\
								G_{3N_z+2} \left( c_f,c_s,h; \Theta \right) 
						\end{bmatrix}}_{G \left( x; \Theta \right)} 
				\end{align*} }
			where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters, $N_{u}$ is the number of control variables.

>>>>

	The fluid density $\rho_f$ can be obtained from the Peng-Robinson equation of state if the temperature and thermodynamic pressure are known along $z$. Variation in fluid density may occur due to pressure or inlet temperature changes. In a non-isothermal case, in Equations \ref{EQ:Velocity_1} and \ref{EQ:Velocity_2} $\rho_f$ is considered the average fluid density along the extraction column.

@@

			The heat balance equation describe the evolution of the enthalpy in the system and it is given by Equation \ref{EQ:Enthalpy_equation}

@@

			The main advantage of this formulation is the presence of term $\partial P / \partial t $, which directly affects the system through the change of thermodynamic pressure.

			If the value of enthalpy $h$ is known from the time evolution of the energy equation, and pressure $P$ is known from measurement, then the temperature $T$ can be reconstructed based on the departure function. The departure function is a mathematical function that characterizes the deviation of a thermodynamic property (enthalpy, entropy, and internal energy) of a real substance from that of an ideal gas at the same temperature and pressure. The departure function is defined as the difference between the value of a thermodynamic property for a real fluid and the corresponding value for an ideal gas at the same temperature and pressure. They are computed by integrating a function that depends on the equation of state and its derivatives. As presented by \citet{Gmehling2019}, for the Peng-Robinson equation of state, the enthalpy departure function is defined by Equation \ref{EQ:Enthalpy_PR}.

@@

			As explained in Chapters \ref{CH:Governing_equations_chapter}, at Low-Mach number conditions, the thermodynamic pressure is nearly constant in space due to the small pressure wave propagation that occurs at the speed of sound. Under such conditions, the term $\partial P/\partial t$ can be approximated by a difference equation, which describes the pressure change in the whole system. The pressure $P$ in the system is considered a state variable, while the pressure in the new time-step $P_{in}$ is considered a control variable.

			{\footnotesize
				\begin{equation}
					\frac{\partial {\color{black}P}}{\partial t} \approx \frac{{\color{black}P_{in}} - {\color{black}P} }{\Delta t}
			\end{equation}}

			Such a simplified equation allows for instantaneous pressure change in the system but does not consider a gradual pressure build-up and the effects of pressure losses. In a real system, the dynamics of pressure change would depend on a pump and a back-pressure regulator.

@@

			{\footnotesize
				\begin{align*}
					&c_f(t = 0, z) = 0  && c_s(t = 0, z) = c_{s0} && h(t = 0, z) = h_0 \\
					&c_f(t,   z=0) = 0  && h(t, z=0) = h_{in}  && \frac{\partial c_f(t,z=L)}{\partial x} = 0 \\
					&\frac{\partial h(t,z=L)}{\partial x} = 0   && c_s(t, z=\{0,L\}) = 0 && y(0) = 0 \quad P(0) = P_0 \\
			\end{align*} }

@@

				{\footnotesize
					\begin{align*} \label{discretization}
						\dot{x} &= \cfrac{d x}{d t} = 
						\begin{bmatrix}
							\cfrac{d c_{f,1}}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{f,N_z}}{d t} \\
							\\ \hline \\
							\cfrac{d c_{s,1}}{d t} 	  \\
							\vdots					  \\
							\cfrac{d c_{s,N_z}}{d t} \\
							\\ \hline \\
							\cfrac{d h_1}{d t} 	  \\
							\vdots 					  \\
							\cfrac{d h_{N_z}}{d t} \\
							\\ \hline \\
							\cfrac{d P}{d t} \\
							\\ \hline \\
							\cfrac{d y}{d t}
						\end{bmatrix}
						=
						\underbrace{\begin{bmatrix}
								G_1 \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\ 
								G_{N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \hline \\ \\
								G_{N_z+1} \left( c_f,c_s,h; \Theta \right)\\ 
								\vdots\\
								G_{2N_z} \left( c_f,c_s,h; \Theta \right)\\ 
								\\ \\ \hline \\ 
								G_{2N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\vdots\\
								G_{3N_z} \left( c_f,c_s,h; \Theta \right) \\ 
								\\ \\ \hline \\
								G_{3N_z+1} \left( c_f,c_s,h; \Theta \right) \\
								\\ \\ \hline \\
								G_{3N_z+2} \left( c_f,c_s,h; \Theta \right) 
						\end{bmatrix}}_{G \left( x; \Theta \right)} 
				\end{align*} }
			where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters, $N_{u}$ is the number of control variables.

### 2024-06-07 Sections/Results_Sensitivity.tex
07.06

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model was based on experiments performed under varying operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow between $3.33-6.67 \cdot 10^{-5}$ kg/s. A sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

>>>>

	This study examines the impact of pressure change on the state space and extraction yield. The process model and parameters are described in {\color{red}article 1}. Calibration of the process model is based on experiments performed under varying operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow between $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis was conducted, assuming the system operates at the midpoint of the validated range: $35^\circ C$, 150 bar and 5 $\cdot 10^{-5}$ kg/s.

### 2024-06-07 Sections/Sensitivity_Analysis.tex
28.05
07.06

	Local derivative-based methods involve taking the total derivative of the state vector $x$ with respect to parameters $p$, where $p$ is a subset of the parameter space $\theta$. The set of derivatives, known as sensitivity equations, is integrated simultaneously with the process model. The sensitivity analysis shows how responsive the solution is for changes in the parameter $p$. As discussed by \citet{Dickinson1976}, the sensitivity analysis can be used to determine the influence of the uncertainty on the solution of the original system. A sensitivity analysis can be used to distinguish sensitive parameters from insensitive ones, which might be helpful for model reduction. Finally, from a control engineering point of view, the sensitivity analysis allows sorting the control variables with respect to the level of effort required to change the model's output.

@@

	{\footnotesize
		\begin{equation}
			\label{SA_def}
			Z(x(t);p) = \cfrac{dx(t)}{dp}
	   \end{equation} }

@@

	{\footnotesize
		\begin{equation} \label{SA_dt} 
			{\dot{Z}}(x(t);p)  = \cfrac{dZ(x(t);p)}{d t} = \cfrac{d}{d t} \left( \cfrac{dx(t)}{dp} \right) = \cfrac{d }{dp} \left( \cfrac{dx(t)}{d t} \right) = \cfrac{d {G}(x(t);p)}{dp} 
	\end{equation} }

	The sensitivity equation can be obtained by applying the definition of the total derivative to Equation \ref{SA_dt}.

	{\footnotesize
		\begin{equation} \label{SA_eq_full}
			\cfrac{dG(x(t);p)}{dp} = \underbrace{ \cfrac{\partial {G}(x(t);p)}{\partial x(t)} }_{{J_x}(x(t);p)} \underbrace{\cfrac{\partial x(t)}{\partial p} }_{S(x(t);p)} + \underbrace{ \cfrac{\partial {G}(x(t);p)}{\partial p} }_{{J_p}(x(t);p)}
	  \end{equation} }

	Equation \ref{SA_eq_full} is solved simultaneously with the process model and consists of three terms: Jacobian ${J_x}(x(t);p)$, sensitivity matrix $S(x(t);p)$ and Jacobian ${J_p}(x(t);p)$. The Jacobian ${J_x}(x(t);p)$ represents the matrix of equations of size $N_x \times N_x$, where each equation ${J_x}(n_x,n_x)$ is the derivative of ${G}_{n_x}(x(t);p)$ with respect to the state variable $x_{n_p}$.

	{\footnotesize
		\begin{align}
			\begingroup % keep the change local
			\setlength\arraycolsep{2pt}
			{J_x}(x(t);p)=\begin{pmatrix}
				\cfrac{\partial {G_{1}}(x(t);p)}{\partial {x_{1}}(t)} & \cfrac{\partial {G_{1}}(x(t);p)}{\partial {x_{2}}(t)} & \cdots & \cfrac{\partial {G_{1}}(x(t);p)}{\partial {x_{N_x}}(t)}\\
				\cfrac{\partial {G_{2}}(x(t);p)}{\partial {x_{1}}(t)} & \cfrac{\partial {G_{2}}(x(t);p)}{\partial {x_{2}}(t)} & \cdots & \cfrac{\partial {G_{2}}(x(t);p)}{\partial {x_{N_x}}(t)}\\
				\vdots & \vdots & \ddots & \vdots\\ 
				\cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {x_{1}}(t)} & \cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {x_{2}}(t)} & \cdots & \cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {x_{N_x}}(t)}\\
			\end{pmatrix}
			\endgroup
	\end{align} }

	The sensitivity matrix $S(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation $S(n_x,n_p)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $p_{n_p}$. Matrix ${J_x}(x(t);p)$ and $S(x(t);p)$ describe indirect influence of $p_{n_p}$ on the model output.

	{\footnotesize
		\begin{equation}
			\begin{split}
				S(x(t);p) & = 
				\begingroup % keep the change local
				\setlength\arraycolsep{2pt}
				\begin{pmatrix}
					\cfrac{\partial {x_{1}}(t)}{\partial {p_{1}}} 	& \cfrac{\partial {x_{1}}(t)}{d {p_{2}}}     & \cdots & \cfrac{d {x_{1}}(t)}{\partial {p_{N_p}}}\\
					\cfrac{\partial {x_{2}}(t)}{\partial {p_{1}}} 	& \cfrac{\partial {x_{2}}(t)}{d {p_{2}}}     & \cdots & \cfrac{d {x_{2}}(t)}{\partial {p_{N_p}}}\\
					\vdots					 	    & \vdots 					   	  & \ddots & \vdots\\
					\cfrac{\partial {x_{N_x}}(t)}{\partial {p_{1}}} 	& \cfrac{\partial {x_{N_x}}(t)}{d {p_{2}}}     & \cdots & \cfrac{\partial {x_{N_x}}(t)}{d {p_{N_p}}}
				\end{pmatrix} 
				\endgroup
			\end{split}
	\end{equation} }

	The Jacobian ${J_p}(x(t);p)$ represents the matrix of equations of size $N_x \times N_p$, where each subequation ${J_p}(n_x,n_p)$ is the partial derivative of the process model equation $F_{n_x}$ with respect to the parameter $p_{n_p}$. ${J_p}(n_x,n_p)$ can be defines direct effect of $p_{n_p}$ on the model output.

	{\footnotesize
		\begin{align}
			{J_p}(x(t);p) & =
			\begingroup % keep the change local
			\setlength\arraycolsep{2pt}
			\begin{pmatrix}
				\cfrac{\partial {G_{1}}(x(t);p)}{\partial {p_{1}}} & \cfrac{\partial {G_{1}}(x(t);p)}{\partial {p_{2}}} & \cdots & \cfrac{\partial {G_{1}}(x(t);p)}{\partial {p_{N_p}}}\\
				\cfrac{\partial {G_{2}}(x(t);p)}{\partial {p_{1}}} & \cfrac{\partial {G_{2}}(x(t);p)}{\partial {p_{2}}} & \cdots & \cfrac{\partial {G_{2}}(x(t);p)}{\partial {p_{N_p}}}\\
				\vdots & \vdots & \ddots & \vdots\\
				\cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {p_{1}}} & \cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {p_{2}}} & \cdots & \cfrac{\partial {G_{N_x}}(x(t);p)}{\partial {p_{N_p}}}
			\end{pmatrix}
			\endgroup
	\end{align}}

	The combined system containing the original set of equations ${G}(x(t);p)$ and sensitivity equations can be formulated as ${\textbf{G}}\left(x(t);p\right)$. The size of ${\textbf{G}}\left(x(t);p\right)$ is equal to $N_s = N_x(N_p + 1)$.

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x(t);p\right) = 
			\begin{bmatrix}
				{G}(x(t);p)\\
				{J_x}(x(t);p)S(x(t);p) + {J_p}(x(t);p)
			\end{bmatrix}
	\end{align} }

@@

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x(t_0);p\right)  &= 
			\begin{bmatrix}
                x(t_0),						               &
				\cfrac{ \text{d}x(t_0) }{ d{p_1} },		   &
				\cdots,					 				   &
				\cfrac{ dx(t_0) }{ d{p_{N_p}} }            
			\end{bmatrix}^\top = \\ &=
            \begin{bmatrix} 
                \quad x_0,	                               &
				\quad ~0,		                           &
			    \quad \cdots,			                   &
				\quad 0 \quad~~
			\end{bmatrix}^\top
	\end{align} }

	The sensitivity analysis of the output function can be performed with respect to parameters $p$. The output function $g(x(t))$ returns ${y}(t)$. By taking a total derivative of ${y}(t)$ with respect to $p$, the new sensitivity equation can be found.

	{\footnotesize
		\begin{equation}
			\cfrac{d {y}(t)}{dp} = \cfrac{d g(x(t))}{dp} = \cfrac{\partial g(x(t))}{\partial x(t)} \cfrac{\partial x(t)}{\partial p} + \cfrac{\partial g(x(t))}{\partial p}
	   \end{equation} }

>>>>

	Local derivative-based methods involve taking the total derivative of the state vector $x$ with respect to the parameter space $\Theta$. The set of derivatives, known as sensitivity equations, is integrated simultaneously with the process model. The sensitivity analysis shows how responsive the solution is for changes in the parameter $\Theta$. As discussed by \citet{Dickinson1976}, the sensitivity analysis can be used to determine the influence of the uncertainty on the solution of the original system. A sensitivity analysis can be used to distinguish sensitive parameters from insensitive ones, which might be helpful for model reduction. Finally, from a control engineering point of view, the sensitivity analysis allows sorting the control variables with respect to the level of effort required to change the model's output.

@@

	{\footnotesize
		\begin{equation}
			\label{SA_def}
			Z(x;\Theta) = \cfrac{dx}{d\Theta}
	   \end{equation} }

@@

	{\footnotesize
		\begin{equation} \label{SA_dt} 
			{\dot{Z}}(x;\Theta)  = \cfrac{dZ(x;\Theta)}{d t} = \cfrac{d}{d t} \left( \cfrac{dx}{d\Theta} \right) = \cfrac{d }{d\Theta} \left( \cfrac{dx}{d t} \right) = \cfrac{d {G}(x;\Theta)}{d\Theta} 
	\end{equation} }

	The sensitivity Equation \ref{SA_eq_full} can be obtained by applying the definition of the total derivative to Equation \ref{SA_dt}.

	{\footnotesize
		\begin{equation} \label{SA_eq_full}
			\cfrac{dG(x;\Theta)}{d\Theta} = \underbrace{ \cfrac{\partial {G}(x;\Theta)}{\partial x} }_{{J_x}(x;\Theta)} \underbrace{\cfrac{\partial x}{\partial \Theta} }_{S(x;\Theta)} + \underbrace{ \cfrac{\partial {G}(x;\Theta)}{\partial \Theta} }_{{J_\Theta}(x;\Theta)}
	  \end{equation} }

	The Jacobian ${J_x}(x;\Theta)$ represents the matrix of equations of size $N_x \times N_x$, where each equation ${J_x}(n_x,n_x)$ is the derivative of ${G}_{n_x}(x;\Theta)$ with respect to the state variable $x_{n_\Theta}$.

	{\footnotesize
		\begin{align}
			\begingroup % keep the change local
			\setlength\arraycolsep{2pt}
			{J_x}(x;\Theta)=\begin{pmatrix}
				\cfrac{\partial {G_{1}}(x;\Theta)}{\partial {x_{1}}} & \cfrac{\partial {G_{1}}(x;\Theta)}{\partial {x_{2}}} & \cdots & \cfrac{\partial {G_{1}}(x;\Theta)}{\partial {x_{N_x}}}\\
				\cfrac{\partial {G_{2}}(x;\Theta)}{\partial {x_{1}}} & \cfrac{\partial {G_{2}}(x;\Theta)}{\partial {x_{2}}} & \cdots & \cfrac{\partial {G_{2}}(x;\Theta)}{\partial {x_{N_x}}}\\
				\vdots & \vdots & \ddots & \vdots\\ 
				\cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {x_{1}}} & \cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {x_{2}}} & \cdots & \cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {x_{N_x}}}\\
			\end{pmatrix}
			\endgroup
	\end{align} }

	The sensitivity matrix $S(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation $S(n_x,n_\Theta)$ is the derivative of the state variable $x_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. Matrix ${J_x}(x;\Theta)$ and $S(x;\Theta)$ describe indirect influence of $\Theta_{n_\Theta}$ on the model output.

	{\footnotesize
		\begin{equation}
			\begin{split}
				S(x;\Theta) & = 
				\begingroup % keep the change local
				\setlength\arraycolsep{2pt}
				\begin{pmatrix}
					\cfrac{\partial {x_{1}}}{\partial {\Theta_{1}}} 	& \cfrac{\partial {x_{1}}}{d {\Theta_{2}}}     & \cdots & \cfrac{d {x_{1}}}{\partial {\Theta_{N_\Theta}}}\\
					\cfrac{\partial {x_{2}}}{\partial {\Theta_{1}}} 	& \cfrac{\partial {x_{2}}}{d {\Theta_{2}}}     & \cdots & \cfrac{d {x_{2}}}{\partial {\Theta_{N_\Theta}}}\\
					\vdots					 	    & \vdots 					   	  & \ddots & \vdots\\
					\cfrac{\partial {x_{N_x}}}{\partial {\Theta_{1}}} 	& \cfrac{\partial {x_{N_x}}}{d {\Theta_{2}}}     & \cdots & \cfrac{\partial {x_{N_x}}}{d {\Theta_{N_\Theta}}}
				\end{pmatrix} 
				\endgroup
			\end{split}
	\end{equation} }

	The Jacobian ${J_\Theta}(x;\Theta)$ represents the matrix of equations of size $N_x \times N_\Theta$, where each subequation ${J_\Theta}(n_x,n_\Theta)$ is the partial derivative of the process model equation $G_{n_x}$ with respect to the parameter $\Theta_{n_\Theta}$. ${J_\Theta}(n_x,n_\Theta)$ defines direct effect of $\Theta_{n_\Theta}$ on the model output.

	{\footnotesize
		\begin{align}
			{J_\Theta}(x;\Theta) & =
			\begingroup % keep the change local
			\setlength\arraycolsep{2pt}
			\begin{pmatrix}
				\cfrac{\partial {G_{1}}(x;\Theta)}{\partial {\Theta_{1}}} & \cfrac{\partial {G_{1}}(x;\Theta)}{\partial {\Theta_{2}}} & \cdots & \cfrac{\partial {G_{1}}(x;\Theta)}{\partial {\Theta_{N_\Theta}}}\\
				\cfrac{\partial {G_{2}}(x;\Theta)}{\partial {\Theta_{1}}} & \cfrac{\partial {G_{2}}(x;\Theta)}{\partial {\Theta_{2}}} & \cdots & \cfrac{\partial {G_{2}}(x;\Theta)}{\partial {\Theta_{N_\Theta}}}\\
				\vdots & \vdots & \ddots & \vdots\\
				\cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {\Theta_{1}}} & \cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {\Theta_{2}}} & \cdots & \cfrac{\partial {G_{N_x}}(x;\Theta)}{\partial {\Theta_{N_\Theta}}}
			\end{pmatrix}
			\endgroup
	\end{align}}

	The combined system containing the original set of equations ${G}(x;\Theta)$ and sensitivity equations can be formulated as ${\textbf{G}}\left(x;\Theta\right)$. The size of ${\textbf{G}}\left(x;\Theta\right)$ is equal to $N_s = N_x(N_\Theta + 1)$.

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x;\Theta\right) = 
			\begin{bmatrix}
				{G}(x;\Theta)\\
				{J_x}(x;\Theta)S(x;\Theta) + {J_\Theta}(x;\Theta)
			\end{bmatrix}
	\end{align} }

@@

	{\footnotesize
		\begin{align}
			{\textbf{G}}\left(x(t_0);\Theta\right)  &= 
			\begin{bmatrix}
                x(t_0),						               &
				\cfrac{ \text{d}x(t_0) }{ d{\Theta_1} },		   &
				\cdots,					 				   &
				\cfrac{ dx(t_0) }{ d{\Theta_{N_\Theta}} }            
			\end{bmatrix}^\top = \\ &=
            \begin{bmatrix} 
                \quad x_0,	                               &
				\quad ~0,		                           &
			    \quad \cdots,			                   &
				\quad 0 \quad~~
			\end{bmatrix}^\top
	\end{align} }

	The sensitivity analysis of the output function can be performed with respect to parameters $\Theta$. The output function $g(x)$ returns ${y}$. By taking a total derivative of ${y}$ with respect to $\Theta$, the new sensitivity equation can be found.

	{\footnotesize
		\begin{equation}
			\cfrac{d {y}}{d\Theta} = \cfrac{d g(x)}{d\Theta} = \cfrac{\partial g(x)}{\partial x} \cfrac{\partial x}{\partial \Theta} + \cfrac{\partial g(x)}{\partial \Theta}
	   \end{equation} }
